---
title: "AdequacyAndLeverage_start"
author: "Dima Mikhaylov"
date: "11/8/2021"
output: html_document
---


```{r}
dataset <- read.table("nfl.txt", header=TRUE)
head(dataset)
```

The variables are:

• y: Games won (14-game season)
• x1: Rushing yards (season)
• x2: Passing yards (season)
• x3: Punting average (yards/punt)
• x4: Field goal percentage (FGs made/FGs attempted)
• x5: Turnover differential (turnovers acquired minus turnovers lost)
• x6: Penalty yards (season)
• x7: Percent rushing (rushing plays/total plays)
• x8: Opponents’ rushing yards (season)
• x9: Opponents’ passing yards (season)


We will continue to regress the number of games won against three predictors: passing
yards, x2, percent rushing, x7, and opponents’ rushing yards in the season, x8.


1. For this first question, you will generate partial regression plots for each of the predictors. As a reminder, a partial regression plot for predictor xk is obtained by:

• Regressing y against the other predictors, x1, · · · , xk−1 ,that are already in the
model and obtaining the residuals, e(y|x1, · · · , xk−1).

• Regressing the predictor in question, xk, against the predictors that are already
in the model and obtaining the residuals, e(xk|x1, · · · , xk−1).

• Plotting the residuals against each other, e(y|x1, · · · , xk−1) against e(xk|x1, · · · , xk−1).


(a) Produce the partial regression plot for x2. Interpret what this partial regression is telling us.

```{r}
lm(y ~ x2, data=dataset)
```

(b) Fit a linear regression for the partial regression plot for x2. Report the estimated
coefficients.



(c) Fit a linear regression for the response against the three predictors. Report the
estimated coefficient for x2 and compare the value with the estimated slope from
the previous part. What do you notice?



(d) Before producing the partial regression plots for x7 and x8, what do you think
will be the values of the estimated coefficients for the linear regression for each of
these plots?



(e) Produce the partial regression plots for x7 and x8. Interpret what both of these
plots are telling us.


2. Using externally studentized residuals, do we have any outliers? What teams are these?



3. Do we have any high leverage data points for this multiple linear regression? What
teams are these?


4. Use DF F IT Si, DF BET ASj,i, and Cook’s distance to check for influential observations. What teams are influential?









