---
title: "M4_examples"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(palmerpenguins)
data(package = 'palmerpenguins')
head(penguins)
```

1. Produce a plot of body mass against flipper length for Gentoo penguins. Write the
estimated linear regression equation.

```{r}
``` {r}
ggplot(penguins[which(penguins$species=='Gentoo'), ], aes(x=flipper_length_mm, y=body_mass_g))+
  geom_point()+
  labs(x="Flipper length (mm)", y="Body mass (grams)",
       title="Scatterplot of Body mass against Flipper length for Gentoo only")
```

```{r}
Gentoo <- na.omit(penguins[which(penguins$species=='Gentoo'),
                   c("flipper_length_mm", "body_mass_g")])

result <- lm(body_mass_g ~ flipper_length_mm, data=Gentoo)
summary(result)
```


## Note: body_mass_g = -6787.281 + 54 * (flipper_length_mm)


2. What is the change in the predicted body mass (in grams) when flipper length increases
by 1mm, for Gentoo penguins? Also report the corresponding 95% confidence interval
for the change in the predicted body mass (in grams) when flipper length increases by
1mm.


## Note: the change in y_hat is 54.623 grams. 95% for the change is given by B1/se(B1):

```{r}
confint(result, level = 0.95)
```


3. Conduct a hypothesis test to determine whether or not there is a linear association
between body mass and flipper length for Gentoo penguins. State the hypotheses,
p-value, and conclusion in context.

## Note: H0: B1 = 0 and Ha: B1 != 0. With p-value for F-statistic of 2.2e-16 we have to reject H0 and conclude that there is not enough evidence to state that the slope is zero and thus there is no relationship mass ~ length.


4. Are your results from parts 2 and 3 consistent? Briefly explain.

#### Note: yes, there are, because from #2 it appears that B1 is in 95% CI and from  #3 it appears that F-test is highly significants as well.

5. Estimate the mean body mass (in grams) for Gentoo penguins with flipper lengths of
200mm. Also report the 95% confidence interval for the mean body mass (in grams)
for Gentoo penguins with flipper lengths of 200mm.

```{r}
-6787.281 + 54 * (200)
```

```{r}
new_data <- data.frame(flipper_length_mm=200)
predict(result, new_data, level=0.95, interval="confidence")
```

6. Report the 95% prediction interval for the body mass (in grams) of a Gentoo penguin
with flipper length of 200mm.

```{r}
new_data <- data.frame(flipper_length_mm=200)
predict(result, new_data, level=0.95, interval="prediction")
```


7. A researcher hypothesizes that for Gentoo penguins, the predicted body mass increases
by more than 50 g for each additional mm in flipper length. Conduct an appropriate
hypothesis test. What is the null and alternative hypotheses, test statistic, and conclusion?

## Note ???
* H0: flipper_length > 50
* Ha: flipper_length =< 50
* test statistic is given by the following 

``` {r}
pt((50-54.623) / 5.028, 121)
```


```{r}
confint(result, level = 0.95)
```







