---
title: "HW1"
author: "Dmitry Mikhaylow"
date: "9/7/2021"
output:
  html_document: default
  pdf_document: default
---

Loading and storing COVID data in R:

``` {r}
Covid <- read.csv('USCovid.csv', header=TRUE)
head(Covid)
```

## 1. Country level analysis

a.    We are interested in the data at the most recent date, June 3 2021. Create a data
frame called `latest` that:
  *   has only rows pertaining to data from June 3 2021:
```{r}
latest <- subset(Covid, date == "2021-06-03")
dim(latest)
```
  * removes rows pertaining to counties that are "Unknown"
```{r}
latest <- subset(latest, latest$county!='Unknown')
dim(latest)
```
  * removes the column `date` and `fips`
```{r}
latest <- subset(latest, select=-c(date, fips))
head(latest)
```

b.    Calculate the death rate (call it `death.rate`) for each county. Report the death
rate as a percent and round to two decimal places. Add death.rate as a new
column to the data frame latest. Display the first 6 rows of the data frame
latest.
```{r}
latest$death.rate = round((latest$deaths / latest$cases) * 100, 2)
head(latest)
```

c. Display the counties with the 10 largest number of cases. Be sure to also display
the number of deaths and death rates in these counties, as well as the state the
counties belong to.
```{r}
head(latest[order(latest$cases, decreasing = TRUE),], 10)
```

d. Display the counties with the 10 largest number of deaths. Be sure to also display
the number of cases and death rates in these counties, as well as the state the
counties belong to.
```{r}
head(latest[order(latest$deaths, decreasing = TRUE),], 10)
```

e. Display the counties with the 10 highest death rates. Be sure to also display the
number of cases and deaths in these counties, as well as the state the counties
belong to. Is there sometime you notice about these counties?

#### Note: death rates are high but the actual number of cases is relatively low:

``` {r}
head(latest[order(latest$death.rate, decreasing = TRUE),], 10)
```

f. Display the counties with the 10 highest death rates among counties with at least
100,000 cases. Be sure to also display the number of cases and deaths in these
counties, as well as the state the counties belong to.
``` {r}
head(latest[order(c(latest$cases, latest$death.rate), decreasing = TRUE),], 10)
```
Another way to do the same:
```{r}
head(latest[order(-latest$cases, latest$death.rate), ], 10)
```

g. Display the number of cases, deaths, death rate for the following counties:
  * Albemarle, Virginia
```{r}
latest[which(latest$county=="Albemarle"), c(3, 4, 5)]
```

  * Charlottesville city, Virginia
```{r}
latest[which(latest$county=="Charlottesville city"), c(3, 4, 5)]
```

## 2. State level analysis

a. We are interested in the data at the most recent date, June 3 2021. Create a data
frame called state.level that:
```{r}
state.level <- subset(Covid, date == "2021-06-03")
state.level <- subset(state.level, select=-c(date, county, fips))
```
  * has 55 rows: 1 for each state, DC, and territory
  * has 3 columns: name of the state, number of cases, number of deaths
```{r}
dim(state.level)
```
  * is ordered alphabetically by name of the state
  * Display the first 6 rows of the data frame state.level.
```{r}
state.level <- aggregate(cbind(cases, deaths)~state, data=state.level, FUN=sum)
head(state.level)
```

b. Calculate the death rate (call it state.rate) for each state. Report the death
rate as a percent and round to two decimal places. Add state.rate as a new
column to the data frame state.level. Display the first 6 rows of the data frame
state.level.
```{r}
state.level$state_rate = round((state.level$deaths/ state.level$cases) * 100, 2)
head(state.level)
```

c. What is the death rate in Virginia?
```{r}
state.level[which(state.level$state=="Virginia"), c(1,4)]
```

d. What is the death rate in Puerto Rico?
```{r}
state.level[which(state.level$state=="Puerto Rico"), ]
```
#### Note: PR has extrememly high state.rate due to low number of reported cases.

e. Which states have the 10 highest death rates?
```{r}
head(state.level[order(-state.level$state_rate), ], 10)
```

f. Which states have the 10 lowest death rates?
```{r}
head(state.level[order(state.level$state_rate), ], 10)
```

g. Export this dataset as a .csv file named stateCovid.csv. 
```{r}
write.csv(state.level, 'stateCovid.csv', row.names = FALSE)
```
