---
title: "ModelValidation"
author: "Dima Mikhaylov"
date: "11/1/2021"
output: html_document
---


For this question, we will use the “nfl.txt” data set that we used in Guided Question Set 6. As a reminder, the data are on NFL team performance from the 1976 season. The variables are:

```{r}
dataset <- read.table("nfl.txt", header=TRUE)
head(dataset)
```

• y: Games won (14-game season)
• x1: Rushing yards (season)
• x2: Passing yards (season)
• x3: Punting average (yards/punt)
• x4: Field goal percentage (FGs made/FGs attempted)
• x5: Turnover differential (turnovers acquired minus turnovers lost)
• x6: Penalty yards (season)
• x7: Percent rushing (rushing plays/total plays)
• x8: Opponents’ rushing yards (season)
• x9: Opponents’ passing yards (season)


1. Use the regsubsets() function from the leaps package to run all possible regressions. Set nbest=2.

```{r}
#install.packages("leaps")
library(leaps)
allregs <- regsubsets(y ~ ., data=dataset, nbest=2)
summary(allregs)
```

2. Identify the model (the predictors and the corresponding estimated coefficients) that is best in terms of

```{r}
names(summary(allregs))
```
(a) Adjusted R2 
```{r}
which.max(summary(allregs)$adjr2)
```
```{r}
coef(allregs, 7)
```



(b) Mallow’s Cp
```{r}
which.min(summary(allregs)$cp) 
```

```{r}
coef(allregs, 2)
```

(c) BIC
```{r}
which.min(summary(allregs)$bic)
```

```{r}
coef(allregs, 2)
```

3. Run forward selection, starting with an intercept-only model. Report the predictors
and the estimated coefficients of the model selected.

```{r}
regnull <- lm(y ~ 1, data=dataset)
regfull <- lm(y ~ ., data=dataset)
step(regnull, scope=list(lower=regnull, upper=regfull), direction = 'forward')
```



4. Run backward elimination, starting with the model with all predictors. Report the predictors and the estimated coefficients of the model selected.

```{r}
step(regfull, scope=list(lower=regnull, upper=regfull), direction = 'backward')
```

x8 + x2 + x7 + x9

5. Run stepwise regression, starting with an intercept-only model. Report the predictors and the estimated coefficients of the model selected.

```{r}
step(regnull, scope=list(lower=regnull, upper=regfull), direction = 'both')
```

6. Write a function that computes the PRESS statistic for a regression model. Hint: the diagonal elements from the hat matrix can be found using the lm.influence() function.


```{r}
#Custom function to calculate PRESS statistics:

PRESS <- function(model) {
    i <- residuals(model)/(1 - lm.influence(model)$hat)
    sum(i^2)
}

```

7. Using the function you wrote in part 6, calculate the PRESS statistic for your regression model with x2,x7,x8 as predictors. Calculate the R2 for this model, 2 and compare this value with its R . What comments can you make about the likely predictive performance of this model?

```{r}
model <- lm(y ~ x2+x7+x8, data=dataset)
summary(model)
```




```{r}
PRESS(model)

```