---
title: "HW7"
author: "Dima Mikhaylow"
date: "10/19/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(faraway)
data(seatpos)
head(seatpos)
```

Predictors from the dataset:
• x1: Age. Age in years
• x2: Weight. Weight in pounds
• x3: HtShoes. Height with shoes in cm
• x4: Ht. Height without shoes in cm
• x5: Seated. Seated height in cm
• x6: Arm. Arm length in cm
• x7: Thigh. Thigh length in cm
• x8: Leg. Lower leg length in cm


1. Fit the full model with all the predictors. Using the summary() function, comment on
the results of the t tests and ANOVA F test from the output.

```{r}
result <- lm(hipcenter ~ ., data=seatpos)
summary(result)
```

2. Briefly explain why, based on your output from part 1, you suspect the model shows
signs of multicollinearity.

#### This is a possibility of multicollinearity becuase all individual predictive variables came out as statistically insignificant, although the overall F-statistic is highly significant.


3. Provide the output for all the pairwise correlations among the predictors. Comment
briefly on the pairwise correlations.

```{r}
pairs(seatpos, lower.panel=NULL)
```

maps

4. Check the variance inflation factors (VIFs). What do these values indicate about
multicollinearity?



5. Looking at the data, we may want to look at the correlations for the variables that
describe length of body parts: HtShoes, Ht, Seated, Arm, Thigh, and Leg. Comment
on the correlations of these six predictors.



6. Since all the six predictors from the previous part are highly correlated, you may decide
to just use one of the predictors and remove the other five from the model. Decide
which predictor out of the six you want to keep, and briefly explain your choice.




7. Based on your choice in part 6, fit a multiple regression with your choice of predictor
to keep, along with the predictors x1 = Age and x2 =Weight. Check the VIFs for this
model. Comment on whether we still have an issue with multicollinearity.




8. Conduct a partial F test to investigate if the predictors you dropped from the full
model were jointly insignificant. Be sure to state a relevant conclusion.




9. Produce a plot of residuals against fitted values for your model from part 7. Based on
the residual plot, comment on the assumptions for the multiple regression model. Also
produce an ACF plot and QQ plot of the residuals, and comment on the plots.



10. Based on your results, write your estimated regression equation from part 7. Also
report the R2 of this model, and compare with the R2 you reported in part 1, for the
model with all predictors. Also comment on the adjusted R2
for both models.

