---
title: "Homework 0"
author: "Dima Mikhaylow"
date: "8/25/2021"
output: html_document
---

## Topic A.2
1. Sampling
  + a. mu, sigma/sqrt(n)
  + b. The Central Limit Theorem
  
2. Manufacturing process with mean=116 and std=4.8
  + a. For one subcomponent of length 118
  ```{r}
  1 - pnorm((118-116)/4.8)
  ```
  + b. For the sample mean of 3 subcomponents:
  ``` {r}
  1 - pnorm((118-116)/(4.8/sqrt(3)))
  ```
  
## Topic A.3 -- Confidence Intervals
3. The goals of constructing confidence intervals:
  * Provide an estimate for the unknown parameter of interest
  * Provide a range of plausible values for the unkonw parameter
  * Provide a measure of uncertainty
  
4. Increasing confidence level ...
  * The multiplier increases in magnitude, thus 
  * The margin of error is also increasing

5. Increasing sample size...
  * Decreases the standard error of sample mean and thus 
  * Margin of error and confidence interval go down (decrease)
  
6. Computing t-multiplier for the mean with given confidence intervals: 
  a. CI = 94%, n = 49
```{r}
qt(1-((1-0.94)/2), 49-1)
```
  b. CI = 86%, n = 82
```{r}
qt(1-((1-0.86)/2), 82-1)
```
  c. CI = 74%, n = 150
```{r}
qt(1-((1-0.74)/2), 150-1)
```


7. Random sample of 100 students had mean GPA of 3.2 with std of 0.2.
  a. Calculate 97% confidence interval for the mean GPA. Firts, t-multiplier is given by qt() function:
```{r}
t = qt((1-(1-0.97)/2), 99)
upper = 3.2 + t*(0.2/sqrt(100))
lower = 3.2 - t*(0.2/sqrt(100))
print(paste("Lower:", lower, "Upper:", upper))
```
  b. Margin of Error (ME) can be found by using t-multiple and standard error or the mean. ME shows how far the sample mean will be away from the true population mean in 97% of random samples of 100 students.
```{r}
ME = t * (0.2/sqrt(100))
print(paste("ME =", ME))
```
  c. No, the mean of 3.25 is not a reasonable estimate, because this value is outside the confidence interval.

## Topic A.4 -- Hypothesis Testing 
8. The goal is to see if an observed estimate is due to random variation in the data or is in fact a true feature of the phenomenon.

9. Hypothesis should be written about an unobservant population parameter. It is evaluated (rejected or failed to reject) by checking observable sample statistics.

10. Drafting hypothesis statements:
  + a. H0(avgMPG == 29 m/g), Ha(avgMPG > 29 m/g), reject H0 => gas mileage improved
  + b. H0(Dimameter == 4 mil), Ha(Diameter != 4 mil), two-sided test as both are bad
  + c. H0(avgTime == 2 hours), Ha(acgTime < 2 hours), travel time improved

11. Larger test statistic gives more evidence against the H0, as it is more unlikely that the observed value is due to a random variation.It shows how dissimilar the sample is from H0. The bigger the magnitude, the more dissimilar the sample is.

## General Questions

## Topic B.2
1. Load cars
  + a. `cars.df` created 
  + b. 32 obs. of 11 variables

2. Load students
  + a. `students.df` created
  + b. 250 obs. of 9 variables
  
3. Libraries installed successfully 

4. Load `faraway` package
  + a. `corn.df` created
  + b. 44 obs. of 2 variables

## Topic B.3
5. The following names are not valied in R:`.2020.Age`, `#Courses.2020`, `1_courses_2020`, and `age(2020)`.

#### Notes on Names
There are only a few rules regarding naming objects in R:
* It must begin with a letter or .
* If it begins with ., then . cannot be followed by a number
* Only letters, numbers, ., and _ can be used 

6. Create a vector
```{r}
practice = c(13, 91, 36, 95, 9, 3, 61, 20, 22, 91)
class(practice)
```

7. Logical tests
```{r}
practice[5] != 10
(practice[1] + practice[2]) < 104
(practice[1] + practice[2]) <= 104
(practice[2]==91) & (practice[9]==22)
(practice[5] < 9) | (practice[6] >= 4)
```

8. Creating and operating on a matrix
``` {r}
Mat.A = matrix(data=c(4, 6, 1, 2, 3, 1), nrow=2, ncol=3)
```

```{r}
names(Mat.A) <- c("Huey", "Dewey", "Loie")
```

```{r}
Mat.A[2, 1]
dim(Mat.A)

```

9. Order of levels in the factor:
```{r}
factor(practice)
```

## Topic B.4
Done!