---
title: "Homework 0"
author: "Dima Mikhaylow"
date: "8/25/2021"
output: html_document
---

## Topic A.2
1. Sampling
  + a. mu, sigma/sqrt(n)
  + b. The Central Limit Theorem
  
2. Manufacturing process with mean=116 and std=4.8
  + a. For one subcomponent of length 118
  ```{r}
  1 - pnorm((118-116)/4.8)
  ```
  + b. For the sample mean of 3 subcomponents:
  ``` {r}
  1 - pnorm((118-116)/(4.8/sqrt(3)))
  ```
  
## Topic A.3 -- Confidence Intervals
3. The goals of constructing confidence intervals:
  * Provide an estimate for the unknown parameter of interest
  * Provide a range of plausible values for the unkown parameter
  * Provide a measure of uncertainty
  
4. Increasing confidence level ...
  * The multiplier increases in magnitude, thus 
  * Increases the UVA carrmargin of error

## Topic A.4 -- Hypothesis Testing 


## General Questions

## Topic B.2
1. Load cars
  + a. `cars.df` created 
  + b. 32 obs. of 11 variables

2. Load students
  + a. `students.df` created
  + b. 250 obs. of 9 variables
  
3. Libraries installed successfully 

4. Load `faraway` package
  + a. `corn.df` created
  + b. 44 obs. of 2 variables

## Topic B.3
5. The following names are not valied in R:`.2020.Age`, `#Courses.2020`, `1_courses_2020`, and `age(2020)`.

#### Notes on Names
There are only a few rules regarding naming objects in R:
* It must begin with a letter or .
* If it begins with ., then . cannot be followed by a number
* Only letters, numbers, ., and _ can be used 

6. Create a vector
```{r}
practice = c(13, 91, 36, 95, 9, 3, 61, 20, 22, 91)
class(practice)
```

7. Logical tests
```{r}
practice[5] != 10
(practice[1] + practice[2]) < 104
(practice[1] + practice[2]) <= 104
(practice[2]==91) & (practice[9]==22)
(practice[5] < 9) | (practice[6] >= 4)
```

8. Creating and operating on a matrix
``` {r}
Mat.A = matrix(data=c(4, 6, 1, 2, 3, 1), nrow=2, ncol=3)
```

```{r}
names(Mat.A) <- c("Huey", "Dewey", "Loie")
```

```{r}
Mat.A[2, 1]
dim(Mat.A)

```

9. Order of levels in the factor:
```{r}
factor(practice)
```

## Topic B.4
Done!