---
title: "Homework2"
author: "Dima Mikhaylov"
date: "9/11/2021"
output: html_document
---

```{r}
library(tidyverse)
```

## Part 1: questions based on `PoliceKillings.csv`
```{r}
PoliceKillings <- read.csv('PoliceKillings.csv', header=TRUE)
```
  
  a. Using the `raceethnicity` variable, create a table and a bar chart that displays
the proportions of victims in each race / ethnic level. Also, use your table and
bar chart in conjunction with the US Census Bureau July 1 2019 estimates to
explain what your data reveal.
```{r}
round(prop.table(table(PoliceKillings$raceethnicity), ) *100, 2)

```

```{r}
raceethnicity_prop <- PoliceKillings %>%
  group_by(raceethnicity)%>%
  summarise(Counts=n())%>%
  mutate(Percent=Counts/nrow(PoliceKillings))

ggplot(raceethnicity_prop, aes(x=raceethnicity, y=Percent))+
  geom_bar(fill="brown3", color='black', stat="identity")+
  theme(axis.text.x=element_text(angle=90),
        plot.title=element_text(hjust=0.5))+
  labs(x="Race/Ethnicity", y="Prop. of killings", title="Dist of variable `raceethnicity`")
```

#### Observation: it seems noteworthy that Black people, although account for only 13% of the US population, have a relatively high proportion of almost 30%. In contrast, White people account for 76% of the population but only 50% proportion of the killings.

b. Convert the variable `age`, the age of the victim, to be numeric, and call this
new variable `age.num`. Use the `is.numeric()` function to confirm that the newly
created variable is numeric (and output the result), and add this new variable to
your data frame.

```{r}
is.numeric(PoliceKillings$age)
```
```{r}
PoliceKillings$age.num <- as.numeric(PoliceKillings$age)
is.numeric(PoliceKillings$age.num)
```

c. Create a density plot of the variable `age.num`. Comment on this density plot.
```{r}
ggplot(PoliceKillings, aes(x=age.num))+
  geom_density(fill="darkorange", color='black')+
  theme(plot.title=element_text(hjust=0.5))+
  labs(x="Victim' Age", y="Density", title="Probability density of variable `age.num`")
```

#### Observation: it seems that the density constantly increases for younger victims, picks at around 35 years and declines till the age of 65 years. The density function is almost flat around zero after 70 years. 

d. Create a visualization to compare the ages of victims across the different race /
ethnicity levels. Comment on the visualization.

```{r}
ggplot(PoliceKillings, aes(x=raceethnicity, y=age.num))+
  geom_violin(fill="darkorange", color='black')+
  theme(plot.title=element_text(hjust=0.5))+
  labs(x="Race/Ethnicity", y="Victim' Age", title="Race specific probability density at different ages")
```

#### Observation: it seems that Asian/Pasific victims tend to be older. In constract, Black and Native American victims tend to be younger. Also, Native American victims appear to be extremely homogeneous in terms of age, possibly due to a very small number of observations 

e. Create a visualization to compare the different causes of death (variable cause)
across the different race / ethnicity levels. Comment on this visualization, specifically on whether the cause of death appears to be independent of the victim’s
race / ethnicity.
```{r}
round(prop.table(table(PoliceKillings$cause))*100, 2)
```

```{r}
ggplot(PoliceKillings, aes(x=raceethnicity, fill=cause))+
  geom_bar(color='black', position="fill")+
  theme(axis.text.x=element_text(angle=90),
        plot.title=element_text(hjust=0.5))+
  labs(x="Race", y="Deaths", title="Causes of death by race")
```

#### Observation: from this plot above it does not appear that the cause of death is dependent of the victim’s race / ethnicity. Another stronger conclusion is probably that all groups suffer from gunshot injuries greatly. Also, surprisingly, there are a number of deaths related to tasers, presumably a non-lethal weapon. 

f. Pick at least two variables from the dataset and create a suitable visualization
of the variables. Comment on what the visualization reveals. You may create
new variables based on existing variables, and describe how you created the new
variables.

```{r}
round(prop.table(table(PoliceKillings$raceethnicity, PoliceKillings$gender), ) *100, 2)
```

```{r}
ggplot(PoliceKillings, aes(x=raceethnicity, y=age.num, fill=as.factor(gender)))+
  geom_boxplot(fcolor='black')+
  theme(axis.text.x=element_text(angle=90),
        plot.title=element_text(hjust=0.5))+
  labs(x="Race", y="Deaths", title="Age and gender of deaths by race")
```

#### Observation, when splitting by gender, first females (from the table above) constitute an absolute minority of the vistims, and second, their age distribution largely mimics males from the same race / ethnicity categories.

## Part 2: questions based on `stateCovid.csv` and `State_pop_election.csv`

The dataset should contain 4 columns:
• the name of the state (55 “states”, the 50 states, plus DC, Puerto Rico, Guam,
Northern Mariana Islands, and the Virgin Islands)
• the number of cases
• the number of deaths
• the death rate, defined as the number of deaths divided by the number of cases
You may realize that when you exported the data file as a .csv file, an extra column
was added to the dataframe. Remove this column.
```{r}
Covid <- read.csv('stateCovid.csv', header=TRUE) #, row.names=1)
dim(Covid)
```
```{r}
head(Covid)
```

a. There is a dataset on Collab, called `State_pop_election.csv`. The data contain
the population of the states from the 2020 census (50 states plus DC and Puerto
Rico), as well as whether the state voted for Biden or Trump in the 2020 presidential elections. Merge these two datasets, `stateCovid.csv` and `State_pop_election.csv`.
Use the `head()` function to display the first 6 rows after merging these two
datasets.

```{r}
Election <- read.csv('State_pop_election.csv', header=TRUE)
Merged <- merge( Covid, Election, by.x="state", by.y = "State")
head(Merged)
```

```{r}
dim(Merged)
```

b. Pick at least two variables from the dataset and create a suitable visualization
of the variables. Comment on what the visualization reveals. You may create
new variables based on existing variables, and decribe how you created the new
variables.

```{r}
ggplot(Merged[!is.na(Merged$Election),], aes(x=Election, y=state_rate))+
  geom_boxplot(fill="darkgreen", color='black')+
  theme(plot.title=element_text(hjust=0.5))+
  labs(x="Voting preferences", y="State death rate", title="State COVID death rate by election results")
```

#### Observation: as shown on the boxplot above, distribution of state death rate does vary by aggregate voting preferences, as measured by which candidate, Biden or Trump, prevailed in the last presidential elections. Noteworthy, that the death rate is somewhat higher and much more variant around the mean in sthe tates supporting Biden. Trump supporting states had lower average and much smaller variance, with some large outliers.


