---
title: "Examples M2"
author: "Dima Mikhaylow"
date: "9/7/2021"
output: html_document
---

## Load the data
```{r}
library(tidyverse)
students <- read.csv("new_students.csv", header = TRUE)
head(students)
```
1. Produce a frequency table of the number of students in each level of `GPA.cat`.
```{r}
students$GPA.cat<-factor(students$GPA.cat, levels=c("low", "moderate", "high"))
table(students$GPA.cat)
```
##Check:
2. Produce a bar chart that summarizes the number of students in each level of GPA.cat.Be sure to remove the bar corresponding to the missing values.(?)
```{r}
ggplot(students[!is.na(students$GPA.cat),], aes(x=GPA.cat))+
  geom_bar(fill="blue", color="black")+
  theme(axis.text.x=element_text(angle=90),
        plot.title=element_text(hjust=0.5))+
  labs(x="GPA category", y="Number of students",
        title="Students by GPA category")
```

3. Create a similar bar chart as you did in part 2, but with proportions instead of counts. Be sure to remove the bar corresponding to the missing values.
```{r}
students_prop <- students[!is.na(students$GPA.cat),]%>%
  group_by(GPA.cat)%>%
  summarise(Counts=n())%>%
  mutate(Percent=Counts/nrow(students[!is.na(students$GPA.cat),]))

ggplot(students_prop, aes(x=GPA.cat, y=Percent))+
  geom_bar(fill="blue", color='black', stat="identity")+
  theme(axis.text.x=element_text(angle=90),
        plot.title=element_text(hjust=0.5))+
  labs(x="GPA category", y="Percent of Students", title="Dist of Categories")
```

4. Produce a frequency table for the number of female and male students and the GPA category.
```{r}
table(students$Gender, students$GPA.cat)
```

5. Produce a table for the percentage of GPA category for each gender. For the percentages, round to 2 decimal places. Comment on the relationship between gender and
GPA category.
```{r}
prop.table(students$Gender, students$GPA.cat)
```

6. Create a bar chart to explore the proportion of GPA categories for female and male students. Be sure to remove the bar corresponding to the missing values.


7. Create a similar bar chart similar to the bar chart in part 6, but split by smoking status. Comment on this bar chart.


8. Create a scatterplot of GPA against the amount of hours spent studying a week. How would you describe the relationship between GPA and amount of time spent studying?
```{r}
ggplot(students, aes(x=GPA, y=StudyHrs))+
  geom_point()
```

9. Edit the scatterplot from part 8 to include information about the number of days the student parties in a month.
```{r}
ggplot(students, aes(x=GPA, y=StudyHrs, size=PartyNum))+
  geom_point(shape=21, alpha=0.5)+
  scale_size(range = c(0.1,5))
```

10. Edit the scatterplot from part 9 to include information about whether the student smokes or not.
```{r}
ggplot(students, aes(x=GPA, y=StudyHrs, size=PartyNum, fill=Smoke))+
  geom_point(shape=21, alpha=0.5)+
  scale_size(range = c(0.1,5))
```

