---
title: "Multiple Regression"
author: "Dima Mikhaylow"
date: "10/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
dataset <- read.table("nfl.txt", header=TRUE)
head(dataset)
```

1. Create a scatterplot matrix and find the correlation between all pairs of variables for
this data set. Answer the following questions based on the output:

```{r}
pairs(dataset, lower.panel=NULL)

```

a. Which predictors appear to be linearly related to the number of wins? Which
predictors do not appear to have a linear relationhsip with the number of wins?

b. Do you notice if any of the predictors are highly correlated with one another? If
so, which ones?

c. What predictors would you first consider to use in a multiplie linear regression?
Briefly explain your choices.




2. Regardless of your answer to the previous question, fit a multiple regression model for
the number of games won against the following three predictors: the team’s passing
yardage, the percentage of rushing plays, and the opponents’ yards rushing. Write the
estimated regression equation.

3. Interpret the estimated coefficient for the predictor x7 in context.

4. A team with x2 = 2000 yards, x7 = 48 percent, and x8 = 2350 yards would like to
estimate the number of games it would win. Also provide a relevant interval for this
estimate with 95% confidence.

5. Using the output for the multiple linear regression model from part 2, answer the
following question from a client: “Is this regression model useful in predicting the
number of wins during the 1976 season?” Be sure to write the null and alternative
hypotheses, state the value of the test statistic, state the p-value, and state a relevant
conclusion. What is the critical value associated with this hypothesis test? Perform
the test at 0.05 significance level.

6. Report the value of the t statistic for the predictor x7. What is the relevant conclusion
from this t statistic? Also report the critical value for this hypothesis test. Perform
the test at 0.05 significance level.

7. Check the regression assumptions by creating a residual plot, an ACF plot of the
residuals, and a QQ plot of the residuals. Comment on these plots.




8. Consider adding another predictor, x1, the team’s rushing yards for the season, to
the model. Interpret the results of the t test for the coefficient of this predictor. A
classmate says: “Since the result of the t test is insignificant, the team’s rushing yards
for the season is not linearly related to the number of wins.” Do you agree with your
classmate’s statement?









