---
title: 'STAT 6021: Project 2'
author: "Connie Cui"
date: "11/26/2021"
output:
  html_document:
    df_print: paged
  html_notebook: default
---


Load in packages
```{r}
library(tidyverse)
library(ggplot2)
library(ggcorrplot)
library(Hmisc)
```
import data:
```{r}
house <- read.csv("house_data.csv")
head(house)
```
find missing data:
```{r}
# list rows of data that have missing values
house[!complete.cases(house),]
```
No missing data.
Check data types of each variable:
```{r}
dim(house) # original set has 22 vars, will be reduced down to 13
```
We will definitely need to change the data type for the date column, and potentially look into creating factors for some of the more ordinal variables.
```{r}
house$date = substr(house$date,1,nchar(house$date)-7)
head(house)
```
Convert date variabe to date type:
```{r}
house$date <- as.Date(house$date, "%Y%m%d")
head(house)
```
Turning view, condition, and grade into ordered factors:
```{r}
house$view <- factor(house$view, ordered = TRUE, levels = c(0, 1, 2, 3, 4))
house$condition <- factor(house$condition, ordered = TRUE, levels = c(1, 2, 3, 4, 5))
house$grade <- factor(house$grade, ordered = TRUE, levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15))
house$waterfront <- factor(house$waterfront, ordered = TRUE, levels = c(0, 1))
```


## Part 2: EDA

```{r}
library(miscset)
library(dplyr)
library(gridExtra)
```

```{r}
names(house)
```

#### Prior to dropping Date and Geotags consider using them for plotting, for example transaction counts by dates?


```{r}
house <- subset(house, select=-c(id,num, date, zipcode, lat, long))
names(house)
```

```{r}
#describe(house)
```


Summary plots:
```{r}
sp1 <- ggplot(house, aes(x=sqft_living, y=price, color=waterfront))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="price",
       title="Possible 'waterfront' interactions")


sp2 <- ggplot(house, aes(x=sqft_living, y=price, color=view))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="price",
       title="Possible 'view' interactions")
  
  
sp3 <- ggplot(house, aes(x=sqft_living, y=price, color=condition))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="price",
       title="Possible 'condition' interactions")
  
  
sp4 <- ggplot(house, aes(x=sqft_living, y=price, color=grade))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="price",
       title="Possible 'grade' interactions")

##produce the 4 density plots in a 2 by 2 matrix
grid.arrange(sp1+scale_color_grey(), 
             sp2+scale_color_grey(), 
             sp3+scale_color_grey(), 
             sp4+scale_color_grey(), 
             ncol = 2, 
             nrow = 2)
```


#### Question: how to deal with indicator (ordinary) varibales in this case? Map to binary classes:

```{r}
cat_vars = c("waterfront", "view", "condition", "grade")
```

```{r}
ggplotGrid(ncol = 2, lapply(c("view", "waterfront", "condition", "grade"),
    function(col) {
        ggplot(house, aes_string(col)) + geom_bar() + coord_flip()
    }))+ scale_color_grey()

```

```{r}
bp1 <- ggplot(house, aes(x=waterfront, y=price))+
geom_boxplot()+
labs(x="waterfront", y="price", title="Price by 'waterfront'")

bp2 <- ggplot(house, aes(x=view, y=price))+
geom_boxplot()+
labs(x="view", y="price", title="Price by 'view'")

bp3 <- ggplot(house, aes(x=condition, y=price))+
geom_boxplot()+
labs(x="condition", y="price", title="Price by 'condition'")

bp4 <- ggplot(house, aes(x=grade, y=price))+
geom_boxplot()+
labs(x="grade", y="price", title="Price by 'grade'")

##produce the 4 plots in a 2 by 2 matrix
grid.arrange(bp1, bp2, bp3, bp4, ncol = 2, nrow = 2)
```

#### Mapping to binary classes and checking distributions and interactions

```{r}
# Changing `view` to 0 for regular view and 1 for every other view
house$view <- factor(ifelse(house$view!=0, 1, 0))
# Changing `condition` to 0 for everything below 3 and 1 otherwise
house$condition <- factor(ifelse(house$condition==1 | house$condition==2 | house$condition==3, 0, 1))
# Changing `grade` to 0 for everything below 7 and 1 otherwise
house$grade <- factor(ifelse(house$grade==1 | house$grade==2 | house$grade==3 |
                      house$grade==4 | house$grade==5 | house$grade==7 , 0, 1))
```


```{r}
ggplotGrid(ncol = 2,
  lapply(c("view", "waterfront", "condition", "grade"),
    function(col) {
        ggplot(house, aes_string(col)) + geom_bar() + coord_flip()
    }))
```

```{r}
bp1 <- ggplot(house, aes(x=waterfront, y=price))+
geom_boxplot()+
labs(x="waterfront", y="price", title="Price by 'waterfront'")

bp2 <- ggplot(house, aes(x=view, y=price))+
geom_boxplot()+
labs(x="view", y="price", title="Price by 'view'")

bp3 <- ggplot(house, aes(x=condition, y=price))+
geom_boxplot()+
labs(x="condition", y="price", title="Price by 'condition'")

bp4 <- ggplot(house, aes(x=grade, y=price))+
geom_boxplot()+
labs(x="grade", y="price", title="Price by 'grade'")

##produce the 4 plots in a 2 by 2 matrix
grid.arrange(bp1, bp2, bp3, bp4, ncol = 2, nrow = 2)
```

#### Checking possible interactions after mapping categorical variables to a larger classes


```{r}
sp1 <- ggplot(house, aes(x=sqft_living, y=price, color=waterfront))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="price",
       title="Remaining 'waterfront' interactions")

sp2 <- ggplot(house, aes(x=sqft_living, y=price, color=view))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="price",
       title="Remaining 'view' interactions")
  
sp3 <- ggplot(house, aes(x=sqft_living, y=price, color=condition))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="price",
       title="Remaining 'condition' interactions")
  
sp4 <- ggplot(house, aes(x=sqft_living, y=price, color=grade))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="price",
       title="Remaining 'grade' interactions")

##produce the 4 plots in a 2 by 2 matrix
grid.arrange(sp1+scale_color_grey(), 
             sp2+scale_color_grey(), 
             sp3+scale_color_grey(), 
             sp4+scale_color_grey(), 
             ncol = 2, 
             nrow = 2)
```


#### Final check, same scatter plots but with log(price) - no visiable interaction with log price.

```{r}

sp1 <- ggplot(house, aes(x=sqft_living, y=log(price), color=waterfront))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="log(price)",
       title="Additive 'waterfront' indicator")

sp2 <- ggplot(house, aes(x=sqft_living, y=log(price), color=view))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  labs(x="sqft_living", 
       y="log(price)",
       title="Additive 'view' indicator")
  
sp3 <- ggplot(house, aes(x=sqft_living, y=log(price), color=condition))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  scale_fill_manual(values=c("red", "blue", "green"))+
  labs(x="sqft_living", 
       y="log(price)",
       title="Additive 'condition' indicator")
  
sp4 <- ggplot(house, aes(x=sqft_living, y=log(price), color=grade))+
  geom_point()+
  geom_smooth(method = "lm", se=FALSE)+
  scale_fill_manual(values=c("red", "blue", "green"))+
  labs(x="sqft_living", 
       y="log(price)",
       title="Additive 'grade' indicator")


##produce the 4 plots in a 2 by 2 matrix
grid.arrange(sp1+scale_color_grey(), 
             sp2+scale_color_grey(), 
             sp3+scale_color_grey(), 
             sp4+scale_color_grey(), 
             ncol = 2, 
             nrow = 2)

```

#### Quantitative pridictors:

```{r}

quant_vars = c("yr_built", "yr_renovated",
               "floors", "bedrooms", "bathrooms", 
               "sqft_living", "sqft_lot", 
               "sqft_above", "sqft_basement", 
               "sqft_living15", "sqft_lot15")

hist.data.frame(house[,quant_vars])

```

#### Checking how many quantitative observations have 0 values

```{r}
colSums(house[,quant_vars] == 0)
```


#### Some homes have no basements (more than 50%) also all homes are expected to have non-zero number of bedrooms (13 zeros) and bathrooms (10 zeroz). Drop sqft_basement column and bathroom/betroom rows with zeros:


```{r}
house <- filter(house, bathrooms != 0, bedrooms != 0)
colSums(house[,quant_vars] == 0)
```


```{r}
house <- subset(house, select=-sqft_basement)
names(house)
```



#### Computing age of the house and removing year_build and year_renovated

```{r}
house$age = ifelse(2021-house$yr_renovated >= 2021-house$yr_built, 2021-house$yr_built, 2021-house$yr_renovated)
head(house)
```

```{r}
house <- subset(house, select=-c(yr_renovated, yr_built))
names(house)
```


#### Modified set of quantitative vars:

```{r}
quant_vars = c("age", "bedrooms", "bathrooms", "floors",
               "sqft_living", "sqft_lot", "sqft_above",  
               "sqft_living15", "sqft_lot15")

hist.data.frame(house[,quant_vars])

```
               
#### Correlations of quantitative vars:

```{r}

grey_shades <- c('#DCDCDC','white',	'#696969')


corr <- round(cor(house[,c("price",quant_vars)]), 1)
ggcorrplot(corr, 
           method = "circle", 
           lab = TRUE,
          # type = "lower", 
           outline.color = "white", 
           ggtheme = ggplot2::theme_gray,
           colors = grey_shades)#c("#6D9EC1", "white", "#E46726"))
```

```{r}
house <- subset(house, select=-sqft_above)
str(house)
```

```{r}
#write_csv(house, 'house_data_cleaned.csv')
```



## TODO: interpreting models https://cran.r-project.org/web/packages/jtools/vignettes/summ.html

## Benchmark regression

```{r}
fit <- lm(log(price) ~ . , data = house)
summary(fit)
```

```{r}
plot(fit)
```


## Regsubsets regression 

```{r}
library(leaps)
library(MASS)
allregs <- regsubsets(log(price) ~ ., data=house, nbest=1)
summary(allregs)
```


```{r}
summary(allregs)$adjr2
```

```{r}
coef(allregs, which.max(summary(allregs)$adjr2))
```


```{r}
coef(allregs, which.min(summary(allregs)$bic))
```

```{r}
best <- lm(log(price) ~ bathrooms+sqft_living+floors+waterfront+view+grade+sqft_living15+age, data=house)
sumary(best)
```



```{r}
regnull <- lm(log(price) ~ 1, data=house)
regfull <- lm(log(price) ~ ., data=house)
step(regfull, scope=list(lower=regnull, upper=regfull), direction = 'backward')
```

  (Intercept)     bathrooms   sqft_living        floors  waterfront.L         view1        grade1 sqft_living15 
 1.180812e+01  9.705813e-02  2.265643e-04  1.235768e-01  3.138767e-01  1.753270e-01  9.176457e-02  1.631818e-04 
          age 
 4.272909e-03 
`

##  Logistic Regression 

```{r}
house$value <- factor(ifelse(house$price > 500000, 1, 0))
prop.table(table(house$value))
```



```{r}
house <- subset(house, select=-price)
names(house)
```


```{r}
##80-20 split
set.seed(1)
sample<-sample.int(nrow(house), floor(.80*nrow(house)), replace = F)
train<-house[sample, ]
test<-house[-sample, ]
head(train)
```

```{r}
prop.table(table(train$value))
```


```{r}
full_model<-glm(value ~ bathrooms+bedrooms+sqft_living+floors+waterfront+view+grade+sqft_living15+age, 
                family="binomial", data=train)
summary(full_model)
```

```{r}
library(ROCR)
# Get predictions based on test holdout set
preds <- predict(full_model, newdata=test, type='response')
rates <- prediction(preds, test$value)
roc_result <- performance(rates, measure = 'tpr', x.measure = 'fpr')
plot(roc_result, title="The ROC curve")+
lines(x = c(0,1), y = c(0, 1), col="red")

```


```{r}
cf <- table(test$value, preds>0.5)
cf
```


```{r}
cat("False positive (type 1 error):", cf[3]/(cf[3]+cf[1]))
```


```{r}
cat("False negative (type 2 error):", cf[2]/(cf[2]+cf[4]))
```

```{r}
cat("Error rate:", (cf[3]+cf[2])/(cf[1]+cf[2]+cf[3]+cf[4]))
```

```{r}
cat("Accuracy:", (cf[1]+cf[4])/(cf[1]+cf[2]+cf[3]+cf[4]))
```

# Check remove sqft_above
# Influetial outliers

```{r}
plot()
```








